\documentclass{article}


\usepackage{natbib}
\PassOptionsToPackage{numbers, compress}{natbib}

\usepackage[utf8]{inputenc} %
\usepackage[T1]{fontenc}    %
\usepackage{hyperref}       %
\usepackage{url}            %
\usepackage{booktabs}       %
\usepackage{amsfonts}       %
\usepackage{nicefrac}       %
\usepackage{microtype}      %
\usepackage{xspace, xcolor}         %

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{cleveref}
\Crefname{equation}{Eq.}{Eqs.} %
\usepackage{bbm}
\usepackage{titletoc}
\usepackage{fullpage}
\usepackage{fancyhdr}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{proposition}[theorem]{Proposition}

\numberwithin{equation}{section}
\numberwithin{theorem}{section}

\input{macro.tex}


\title{Beyond NTK with Vanilla Gradient Descent: A Mean-Field Analysis of Neural Networks with Polynomial Width, Samples, and Time}




\author{%
Arvind Mahankali\thanks{Equal Contribution} \\
Stanford University\\
\texttt{amahanka@stanford.edu} \\
\and
Jeff Z. Haochen\footnotemark[1] \\
Stanford University\\
\texttt{jhaochen@stanford.edu} \\
\and
Kefan Dong \\
Stanford University\\
\texttt{kefandong@stanford.edu} \\
\and
Margalit Glasgow \\
Stanford University\\
\texttt{mglasgow@stanford.edu} \\
\and
Tengyu Ma \\
Stanford University \\
\texttt{tengyuma@stanford.edu}
}


\begin{document}
	
	
\maketitle
	


 
	\begin{abstract}
		Despite recent theoretical progress on the non-convex optimization of two-layer neural networks, it is still an open question whether gradient descent on neural networks without unnatural modifications can achieve better sample complexity than kernel methods. This paper provides a clean mean-field analysis of projected gradient flow on polynomial-width two-layer neural networks. 
		Different from prior works, our analysis does not require unnatural modifications of the optimization algorithm. We prove that with sample size $n = O(d^{3.1})$ where $d$ is the dimension of the inputs, the network trained with projected gradient flow converges in $\poly(d)$ time to a non-trivial error that is not achievable by kernel methods using $n \ll  d^4$ samples, hence demonstrating a clear separation between unmodified gradient descent and NTK. As a corollary, we show that projected gradient descent with a positive learning rate and a polynomial number of iterations converges to low error with the same sample complexity.
	\end{abstract}




\input{introduction}
\input{prelim}
\input{setup}

\input{population}
\input{sample_complexity_main_body}
\input{additional-related-work}

\section{Conclusion}

In this paper, we prove a clear sample complexity separation between vanilla gradient flow and kernel methods with any inner product kernel, including NTK. Our work leads to several directions for future research. The first question is to generalize our results to the ReLU activation. Another question is whether gradient descent can achieve less than $d^3$ sample complexity in our setting. The work of \citet{arous2021online} shows that if the population loss has information exponent $2$, then a sample complexity of $d$ (up to logarithmic factors) can be achieved with a single-neuron student network --- it is an open question if a similar sample complexity could be obtained in our setting. A final open question is whether two-layer neural networks can be shown to attain arbitrarily small generalization error $\epsilon$ --- one limitation of our analysis is that we require $\epsilon\ge \poly(1/\log \log d)$.

\section*{Acknowledgments}

The authors would like to thank Zhiyuan Li for helpful discussions. The authors would like to thank the support of NSF IIS 2045685, NSF Grant CCF-1844628 and a Sloan Research Fellowship.

\bibliographystyle{plainnat}
\bibliography{all,new,sample}


\newpage
\appendix
\section*{List of Appendices}
\startcontents[sections]
\printcontents[sections]{l}{1}{\setcounter{tocdepth}{3}}
\newpage

\input{appendix-spherical-harmonics}
\input{population-loss-infinite-width}
\input{finite-width-sample-complexity}
\input{projected-GD-small-LR}
\input{appendix-population-dynamics}
\input{appendix-lower-bound}
\input{toolbox}


\end{document}
